---
title: "Subsidy–Atress Models"
author: "Tasks 4 Group" 
date: "April 2025"
output: html_document
extra_dependencies: "subfig"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Modeling the subsidy–stress relationship
Growth or performance depends on internal metal concentrations $Q$. 
The Droop equation(Droop 1983) can be used for limited growth to capture reduced performance when metals are limiting:
$$\text{Subsidy growth} =\mu\left(1-\frac{q}{Q}\right)$$
Log-logistic dose response equations (Ritz 2010) can be used to capture reduced performance when metal concentrations are high:
$$\text{Stress growth} =\mu\left(c+\frac{d-c}{1+\left(\frac{Q}{a}\right)^b}\right)$$
Combining these functions together with a multiplicative approach yields a subsidy-stress growth 
$$G(Q) =\mu\left(1-\frac{q}{Q}\right)\left(c+\frac{d-c}{1+\left(\frac{Q}{a}\right)^b}\right)$$

```{r initializeParams, echo=FALSE}
#Initialization: load parameter values:
    mu=1.2   #max growth rate
    D=0.2    #death rate
    ke=0.1   #elimination rate
    ku=0.1   #uptake rate
    q=0.005  #min Q required for growth 
    a=0.17   #log-logistic dose-response parameters 
    b=15
    c=0
    d=1
    Me=1  #environmental load
  
 #set parameter list:
parameter_values<-c(mu=mu, D=D, ke=ke, ku=ku, q=q, a=a, b=b, c=c, d=d, Me=Me)

Qvec<-seq(0, 0.25, length.out = 100)
Growth1<-mu*(1-q/Qvec);
Growth2=mu*(c+(d-c)/(1+(Qvec/a)^b)); 
Growth3=mu*(1-q/Qvec)*(c+(d-c)/(1+(Qvec/a)^b));
```

```{r PlotGrowth, fig.cap='Figure 1', fig.subcap=c('(a)', '(b)', '(c)'), fig.ncol=3, out.width="33%", fig.align="center", echo=FALSE}
plot(Qvec, Growth1, type="l", ylim=c(0,1.2),lwd=3, ylab="Subsidy Growth", xlab="Q", cex.axis=1.5, cex.lab=1.5)
plot(Qvec, Growth2, type="l", ylim=c(0,1.2),lwd=3, ylab="Stress Growth", xlab="Q", cex.axis=1.5, cex.lab=1.5)
plot(Qvec, Growth3, type="l", ylim=c(0,1.2), lwd=3,ylab="Subsidy-Stress Growth", xlab="Q",cex.axis=1.5, cex.lab=1.5)
```

We are using systems of ordinary differential equations to model uptake, elimination, and growth dilution of metal quotas. The base model for a simple primary producer 

$$
\begin{align}
\frac{dC}{dt} &=G(Q)C-DC \\
\frac{dQ}{dt} &=k_uM_\text{env}-K_eQ-G(Q)Q
\end{align}
$$
```{r SolveODE, echo = FALSE}
#install.packages("deSolve") #required package for ODE solver:
library(deSolve) 

#Model Equations: 
Model_SubsidyStress <- function(t, variables, parameters) 
{with( as.list(c(variables,parameters)),  #lets us access variables and parameters by name 
       { 
       
       G = mu*(1-q/Q)*(c+(d-c)/(1+(Q/a)^b)); #Growth 
       D = D #Death 
       U = ku*Me #Uptake
       E = ke #Elimination
       #System of ODES:
       dC = G*C-D*C
       dQ = U-E*Q-G*Q
       return(list(c(dC,dQ)))
       }) 
} #End Model Equations


  #Initialization: Set time:
  tend=120 #Days for simulation
  time_values<-seq(0,tend)
  
#Initialization: Set initial conditions:  
  initial_values <- c(C=1,Q=q)

#Solve Model 
Model_solutions <- ode(
  y = initial_values,
  times = time_values,
  func = Model_SubsidyStress,
  parms = parameter_values 
)

Model_solutions<-as.data.frame(Model_solutions) #convert output to dataframe so we can use with()
```

```{r PlotSims, fig.cap='Figure 2', fig.subcap=c('(a)', '(b)'), fig.ncol=1, out.width="50%", fig.align="center", echo=FALSE}
with(Model_solutions, {plot(time, Q, type = "l", xlab = "time (days)", ylab = "Q", cex.axis=1.5, cex.lab=1.5,lwd=3)})
#with(Model_solutions, {plot(time, C, type = "l", xlab = "time (days)", ylab = "Population Density",cex.axis=1.5, cex.lab=1.5,lwd=3, ) })
```

Looping over $M_\text{env}$ and solving for the steady-state values of $Q$ yields 

```{r LoopMenv, echo=FALSE}
#Define Menv vector to loop over ranges of values:
Me_vec<-seq(0, 2, length.out = 100)

#Define Matrix to store results 
Eq_solutions<-matrix(0, nrow=length(Me_vec), ncol=2) #2 state variables
colnames(Eq_solutions) <- c("C", "Q")
C_eq<-vector("numeric" , length(Me_vec))
Q_eq<-vector("numeric" , length(Me_vec))

for(i in 1:length(Me_vec)){
    #replace beta parameter value 
    Me<-Me_vec[i]
    #reset parameter list:
    parameter_values<-c(mu=mu, D=D, ke=ke, ku=ku, q=q, a=a, b=b, c=c, d=d, Me=Me)
    
    #Solve Model 
    Model_solutions <- ode(
      y = initial_values,
      times = time_values,
      func = Model_SubsidyStress,
      parms = parameter_values 
    )
    #Store results of last values in row of matrix 
    Eq_solutions[i, ]<-Model_solutions[tend, 2:3]
   
    Model_solutions_df<-as.data.frame(Model_solutions) #convert output to dataframe for max function 
    C_eq[i]<-max(Model_solutions_df$C)
    Q_eq[i]<-max(Model_solutions_df$Q)
  }

Eq_solutions<-as.data.frame(Eq_solutions) #convert output to dataframe so we can use with()
```

```{r PlotEq, fig.cap='Figure 3', fig.subcap=c('(a)', '(b)'), fig.ncol=1, out.width="50%", fig.align="center", echo=FALSE}
with(Eq_solutions, {plot(Me_vec, Q_eq, xlab = "Environmental Metal Load", ylab = "Q",cex.axis=1.5, cex.lab=1.5,lwd=3)})
with(Eq_solutions, {plot(Me_vec, C_eq, xlab = "Environmental Metal Load", ylab = "C",cex.axis=1.5, cex.lab=1.5,lwd=3)})
```



